---
title: "differences between weekends and weekdays"
author: "me"
date: "7/12/2020"
output: html_document
---
i have unziped the files,then read them and then took the number of steps taken per day
```{r}
activity<-read.csv("activity.csv")
activity_steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
print(activity_steps_day)
``` 
then i take the histogram
```{r}
hist(activity_steps_day$steps, xlab = "Steps per Day", main = "Total number of steps taken per day", col = "red")
```
then i calculate the mean and median
```{r}
mean_steps <- mean(activity_steps_day$steps)
median_steps <- median(activity_steps_day$steps)
print(mean_steps)
print(median_steps)
```
first i calculate the mean of all days and then i make the plot
```{r}
activity<-read.csv("activity.csv")
activity_steps_mean <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
plot(activity_steps_mean$interval, activity_steps_mean$steps, type = "l", col = "tan3", xlab = "Intervals", ylab = "Total steps per interval", main = "Number of steps per interval (averaged) (NA removed)")
```
then to find the maximum steps
```{r}
max_steps <-max(activity_steps_mean$steps)
max_interval <- activity_steps_mean$interval[which(activity_steps_mean$steps == max_steps)]
max_steps <- round(max_steps, digits = 2)
print(max_steps)
print(max_interval)
```
first i am gonna count the total missing values
```{r}
activity<-read.csv("activity.csv")
sum(is.na(activity))
```

creating the new dataset
```{r}
MeanStepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
activity_NAs <- activity[is.na(activity$steps),]
activity_non_NAs <- activity[!is.na(activity$steps),]
activity_NAs$steps <- as.factor(activity_NAs$interval)
levels(activity_NAs$steps) <- MeanStepsPerInterval
levels(activity_NAs$steps) <- round(as.numeric(levels(activity_NAs$steps)))
activity_NAs$steps <- as.integer(as.vector(activity_NAs$steps))
imputed_activity <- rbind(activity_NAs, activity_non_NAs)
```
now the histogram
```{r}
par(mfrow = c(1,2))
activity_steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
hist(activity_steps_day$steps, xlab = "Steps per Day", main = "NAs REMOVED - Total steps/day", col = "green")
imp_activity_steps_day <- aggregate(steps ~ date, data = imputed_activity, FUN = sum, na.rm = TRUE)
hist(imp_activity_steps_day$steps, xlab = "Steps per Day", main = "NAs IMPUTED - Total steps/day", col = "red")
```
now we calculate the new mean and median and compare them to the old ones
```{r}
imp_mean_steps <- mean(imp_activity_steps_day$steps)
imp_median_steps <- median(imp_activity_steps_day$steps)
mean_steps <- mean(activity_steps_day$steps)
median_steps <- median(activity_steps_day$steps)
results_mean_median <- data.frame(c(mean_steps, median_steps), c(imp_mean_steps, imp_median_steps))
colnames(results_mean_median) <- c("NA removed", "Imputed NA values")
rownames(results_mean_median) <- c("mean", "median")
library(xtable)
xtab<-xtable(results_mean_median)
print(xtab,type="html")
```

```{r}
activity<-read.csv("activity.csv")
MeanStepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)

activity_NAs <- activity[is.na(activity$steps),]
activity_non_NAs <- activity[!is.na(activity$steps),]

activity_NAs$steps <- as.factor(activity_NAs$interval)
levels(activity_NAs$steps) <- MeanStepsPerInterval

levels(activity_NAs$steps) <- round(as.numeric(levels(activity_NAs$steps)))
activity_NAs$steps <- as.integer(as.vector(activity_NAs$steps))

imputed_activity <- rbind(activity_NAs, activity_non_NAs)
imputed_activity$dayType <- ifelse(weekdays(as.Date(imputed_activity$date)) == "Samstag" | weekdays(as.Date(imputed_activity$date)) == "Sonntag", "weekend", "weekday")
imputed_activity$dayType <- factor(imputed_activity$dayType)
steps_interval_dayType <- aggregate(steps ~ interval + dayType, data = imputed_activity, FUN = mean)
head(steps_interval_dayType)
names(steps_interval_dayType) <- c("interval", "day_type", "mean_steps")
library(ggplot2)
plot <- ggplot(steps_interval_dayType, aes(interval, mean_steps))
plot + geom_line(color = "red") + facet_grid(day_type~.) + labs(x = "Intervals", y = "Average Steps", title = "Activity Patterns")